---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# retention.helpers

<!-- badges: start -->
<!-- badges: end -->

The goal of retention.helpers is to provide a few functions to help working with the CSU Retention Team's data set, within the R environment.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("benwhicks/retention.helpers")
```

You should also have the tidyverse installed and loaded to get the most out of the package:

``` r
# install.packages("tidyverse")
library(tidyverse)
```

The best way to interface with the data is through the `retention.data` package. This package is not shared publicly, so you will need to obtain a copy of the R Package itself and build the package yourself. This is done in RStudio using `Ctrl-Shift-B` or by going to the _Build_ pane (usually in the top right along with _Environment_) and selecting _Install and Restart_. 

## Examples

### Loading Data

This package is designed to work _with_ the retention data set. This can be loaded two ways:

You can load from a folder containing the .rda files directly:

``` {r example-loading-data-rda-files, eval = FALSE}
# ============================================= #
# Loading via a folder of the .rda files
# ============================================= #
library(tidyverse)
folder_of_rda_files <- file.path('~', 'path', 'to', 'folder') # Change this to where you have them stored
rda_files <- list.files(folder_of_rda_files, 
                        pattern = ".rda", 
                        full.names = TRUE) # This grabs only rda files in folder_of_rda_files
# Loading the data into your environment
rda_files %>% 
  map(load)
```

Or you can load from the `retention.data` package:

```{r example-loading-data-from-package, eval=FALSE}
# ============================================= #
# Loading data via the retention.data package
# ============================================= #
library(retention.data)
```

The benefit of using the package version of the data is that it comes with in built documentation, using R's usual help syntax, `?`.

``` {r example-help, eval = FALSE}
?student_progress
```

### Wrangling

``` {r creating-dummy-data, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
student_ids <- tibble(
  id = c("1", "2A", "3"),
  user_id = c("egalois", "enoether", "fnightingale"),
  email = c("egalois@email.com", "enoether@email.com", "fnight@email.com"),
  firstname = c("Evariste", "Emmy", "Florence"),
  lastname = c("Galois", "Noether", "Nightingale")
)

student_demographics <- tibble(
  id = c("1", "2A", "3"),
  firstname = c("Evariste", "Emmy", "Florence"),
  lastname = c("Galois", "Noether", "Nightingale"),
  yob_approx = c(1811,1882,1820),
  gender = c("Male", "Female", "Female"),
  domesticity = rep("Domestic", 3),
  ses = c("Low SES", "Med SES", "High SES") %>% 
    fct_relevel("Low SES", "Med SES", "High SES")
)

student_progress <- tibble(
  id = c("1", "2A", "2A", "3", "3"),
  session = c(189030, 189030, 189060, 189030, 181960),
  course = c(rep("Applied Mathematics", 5)),
  course_level = c(rep("Undergrad", 5)),
  campus = c("Paris", "Bavaria", "Bavaria", "Hampshire", "Hampshire"),
  course_enrolment = rep("Active Student", 5),
  commencing = c("Commencing", "Commencing", "Not Commencing", "Commencing", "Not Commencing"),
  attendance_type = c("Distance", rep("On Campus", 4)),
  progress_rate = c(0.5, 1, 0.8, 1, 1)
)

academic <- tibble(
  id = c(rep("1", 4), 
         rep("2A", 8),
         rep("3", 8)),
  session = c(rep(189030, 4), 
              rep(189030, 3), rep(189060, 5), 
              rep(189030, 4), rep(189060, 4)),
  subject = c("MTH100", "MTH110", "MTH120", "PHL100",
              "MTH100", "MTH110", "MTH120", "MTH105", "MTH115", "MTH125", "PHL105", "MTH302",
              "MTH100", "MTH110", "MTH120", "PHL100", "MTH105", "MTH115", "MTH125", "PHL105"),
  grade = c("AW", "CR", "FL", "AW", 
            "HD", "DI", "DI", "CR", "PS", "DI", "PS", "FW",
            "PS", "CR", "CR", "CR", "CR", "PS", "DI", "DI") %>% 
    fct_relevel("AW", "FW", "FL", "PS", "CR", "DI", "HD")
)

flags <- tibble(
  id = c("1", "1", "1", "2A"),
  subject = c("MTH120", "PHL100", NA_character_, "MTH302"),
  session = c(189030, 189030, 189060, 189060),
  offering = c("MTH120_189030_P_D", "PHL100_189030_P_D", NA_character_, "MTH302_189060_B_I"),
  campaign = c("pre census", "pre census", "former fail", "pre census"),
  week = c(3, 3, 1, 3),
  concern = c("low activity", "non submission", "prior performance", "non submission")
)
```

This package has a handful of functions for common data wrangling-munging-tidying tasks using the retention data. Most involve adding a derived variable from an existing data frame. For example if `offering` is in the data frame (which comes in the form `ABC123_201990_W_D`) then we can extract the subject `ABC123`, session `201990` or year `2019`. 

### Adding variables

``` {r example-adding-variables}
# creating a data frame with offerings
library(retention.helpers)
offs <- tibble(offering = c("MTH120_189030_P_D", "PHL100_189030_P_D"))
offs

# adding the subject and year
offs %>% 
  add_subject_from_offering() %>% 
  add_year_from_offering()
```

### Joining 

``` {r example-joining}
# Joining demographics and academic
student_demographics %>% 
  inner_join(academic, by = "id") %>% 
  head()

# Joining progress and academic and flags
flags %>% 
  inner_join(academic, by = c("id", "subject", "session"))
```

### Visualising

``` {r example-visualising}
# Trying to visualise grade distributions of flagged vs non flagged
academic %>% 
  left_join(
    flags %>% 
      filter(campaign == "pre census") %>% # only looking at pre census flags
      select(id, subject, session) %>% # only selecting what is required for joining with academic
      mutate(flagged = "At risk"), # left join as we want all academic, which creates NA's if no match
    by = c("id", "subject", "session")) %>%  
  mutate(flagged = replace_na(flagged, "Not at risk")) %>% # NA's did not appear in the flag table
  ggplot(aes(grade)) +
  stat_count(aes(fill = flagged)) +
  facet_grid(flagged ~ session) +
  theme_minimal() 
```

## Data Model

``` {r data-model, echo = FALSE}
# TODO: Copy the data_model.yml file across from the retention.data package to the root of this package whenever this is updated.
library(datamodelr)
dm <- dm_read_yaml('data_model.yml')
gp <- dm_create_graph(dm, rankdir = "RL")
dm_render_graph(gp)
```
