---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
library(retention.helpers)
```

# retention.helpers

<!-- badges: start -->
<!-- badges: end -->

The goal of retention.helpers is to provide a few functions to help working with the CSU Retention Team's data set, within the R environment.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("benwhicks/retention.helpers")
```

You should also have the tidyverse installed and loaded to get the most out of the package:

``` r
# install.packages("tidyverse")
library(tidyverse)
```

The best way to interface with the data is through the `data.csu.retention` package. This package is not shared publicly, so you will need to obtain a copy of the R Package itself and build the package yourself. This is done in RStudio using `Ctrl-Shift-B` or by going to the _Build_ pane (usually in the top right along with _Environment_) and selecting _Install and Restart_. 

## Overview of functions

Most of the functions are split into groups based on their use and prefix. The main families are `add_` and `read_`.

### `add_*`

The `add_` family of functions create new columns in a data frame based on data *in that table only*. This means they are the most broadly useful as the do not require `retention.data` loaded. For example if `offering` is in the data frame (in the form `ABC123_201990_W_D`) then we can extract the subject `ABC123`, session `201990` or year `2019`. 

<details><summary>See an example of using the `add_*` functions</summary>
``` {r add-example}
# Given some data with offering
dat_with_offering <- tibble(
  offering = c("MTH100_189030_P_D", "PHL105_189060_B_I")
)

dat_with_offering

# Add in subject and year from offering
dat_with_offering %>% 
  add_subject_from_offering() %>% 
  add_year_from_offering()
```
</details>

### `read_*`

The `read_*` style functions operate in a similar fashion to `read_csv` in the tidyverse, expect they are designed to read particular commonly used files and clean them in the process. 


## Loading the Data

This package is designed to work _with_ the retention data set. 

```{r example-loading-data-from-package, eval=FALSE}
# ============================================= #
# Loading data via the retention.data package
# ============================================= #
library(data.csu.retention)
```

### Toy data

There is also a sample of this data available (simulated and resampled from the original data, so it does not represent any actual student data). These tables are included in this package and all start with the prefix `toy_`.

``` {r}
glimpse(toy_academic)
```

## Data Model

...coming soon

## Vignettes

See 'Articles' for now - until I work out how to link them properly.

