---
title: "Visualisation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(retention.helpers)
library(tidyverse)
```

Renaming relevant toy data to names from the `data.csu.retention` data package.

``` {r}
academic <- toy_academic_for_interventions
flags <- toy_flags
interventions <- toy_interventions
```

## Visualising a grade distribution

We may like to display grade distributions of the academic data. The data is already set up in factor levels for the major grades, and there are also colour palettes to use; `csu_colours`, `csu_colours_dark` and `csu_colours_light`.

### Basic grade distribution in Charles Sturt colours

``` {r}
academic |> 
    add_grade_helpers() |> 
    filter(grade_substantive) |> 
    ggplot(aes(x = grade, fill = grade_success)) +
    geom_histogram(stat = "count") +
    theme_minimal() +
    scale_fill_manual(values = csu_colours, guide = NULL)
```

### Using facets to compare grade distributions

We may also want to compare grade distributions of different groups. For instance, the table `flags` and `interventions` include data on students *flagged* for some concern in the various campaigns, and the immediate results of the intervention. The toy data included in this package is a simulated sample based on a pre-census campaign were students are called if they are flagged as 'at-risk'. Some answer the phone call, and some do not. 

``` {r}
academic |> 
    inner_join( # only including subjects from the flags list
        flags |> 
            distinct(subject)
    ) |> 
    left_join(
        flags |> 
            mutate(flagged = TRUE)
    ) |> 
    mutate(flagged = replace_na(flagged, FALSE)) |> 
    mutate(grp = if_else(flagged, "Low engagement", "Engaged")) |> 
    add_grade_helpers() |> 
    filter(grade_substantive) |> 
    ggplot(aes(x = grade, fill = grade_success)) +
    geom_histogram(stat = "count") +
    theme_minimal() +
    scale_fill_manual(values = csu_colours, guide = NULL) +
    facet_wrap(~grp, scales = "free_y")
    
```

``` {r}
academic |> 
    inner_join( # only including subjects from the flags list
        flags |> 
            distinct(subject)
    ) |> 
    left_join(
        flags |> 
            mutate(flagged = TRUE)
    ) |> 
    mutate(flagged = replace_na(flagged, FALSE)) |> 
    left_join(interventions) |> 
    mutate(grp = case_when(
        !flagged ~ "Engaged",
        intervention_result == "dialogue" ~ "Spoke with disengaged student",
        intervention_result == "no dialogue" ~ "Disengaged student did not answer") |> 
            fct_relevel("Engaged", "Spoke with disengaged student")) |> 
    add_grade_helpers() |> 
    filter(grade_substantive) |> 
    ggplot(aes(x = grade, fill = grp, alpha = grade_success)) +
    geom_histogram(stat = "count") +
    theme_minimal() +
    scale_fill_manual(values = csu_colours, guide = NULL) +
    scale_alpha_discrete(range = c(1, 0.5), guide = NULL) +
    facet_wrap(~grp, scales = "free_y")
    
```

## Aggregating grades for visualisation

The interventions (for the campaign in the toy data at least) are performed at a **student-session** level, but grades are reported at a **student-session-subject** level. As such it can be useful to aggregate grades from **student-session-subject** to **student-session** level. 

``` {r}
# aggregating at student-session progress rate
academic_summary_1 <- 
    academic |> 
    mutate(session = "Session 1") |> # normally session would be included
    group_by(id, session) |>
    add_grade_helpers() |> 
    summarise(
        progress_rate = sum(grade_success) / sum(grade_substantive),
        gpa = gpa(grade)
    ) |> 
    ungroup()
    
academic_summary_1 |> 
    left_join(interventions) |>
    mutate(grp = case_when(
        is.na(intervention_result)  ~ "Engaged",
        intervention_result == "dialogue" ~ "Spoke with disengaged student",
        intervention_result == "no dialogue" ~ "Disengaged student did not answer") |> 
            fct_relevel("Engaged", "Spoke with disengaged student"))|> 
    ggplot(aes(x = gpa, colour = grp)) +
    stat_ecdf(size = 2, alpha = 0.8) +
    scale_y_continuous(labels = scales::percent_format(),
                       name = "Cumulative distribution") +
    theme_minimal() +
    scale_color_manual(values = csu_colours)
```


``` {r}
# aggregating by the custom `summarise_academic` function
academic_summary_2 <- 
    academic |> 
    mutate(session = "Session 1") |> # normally session would be included
    group_by(id, session) |>
    summarise_academic()  
    
glimpse(academic_summary_2)

academic_summary_2 |> 
    left_join(interventions) |>
    mutate(grp = case_when(
        is.na(intervention_result)  ~ "Engaged",
        intervention_result == "dialogue" ~ "Spoke with disengaged student",
        intervention_result == "no dialogue" ~ "Disengaged student did not answer") |> 
            fct_relevel("Engaged", "Spoke with disengaged student"))|> 
    ggplot(aes(x = 1, fill = result)) +
    geom_bar(stat = "count", position = "stack") +
    facet_wrap(~grp, scales = "free_y") +
    theme_minimal() +
    theme(axis.text.x = element_blank(), axis.title.x = element_blank()) +
    scale_fill_manual(values = csu_colours)

```
