---
title: "Working-with-enrolments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working-with-enrolments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(retention.helpers)
library(tidyverse)
```

## Overview

Enrolments come in two levels at Charles Sturt University; course and subject. Course enrolments refer to the students' enrolment in a particular program of study, such as a Bachelor of Nursing. These are stored in the `student_course` table. Subject level enrolments are stored in the `enrolments` table, and correspond to the enrolment records for units of study in a particular session. These should match to the `academic` table that stores the grades for each subject taken (if the student remained past census date). 

### Quirks

The `enrolments` tables also includes other pseudo-subjects, that are place holders for a Leave of Absence. These 'subjects' have the code **XLV0000** and should not be treated like a normal subject. Subjects beginning with the code **SSS...** might also be needed to be treated differently, as they are free subjects and exhibit very different behaviour from students. 

## Counting enrolments

For this example, we set up a toy version of the `student_course` and `enrolments` tables. We rename them to match the names used in the (private) `data.csu.retention` package.

``` {r}
student_course <- retention.helpers::toy_student_course
enrolments <- retention.helpers::toy_enrolments

glimpse(student_course)
glimpse(enrolments)
```

We might want to look at the active enrolments in a particular session.

``` {r}
# gets all enrolments, and counts by student
enrolments |> 
    filter(session == 202230) |> 
    count(id, sort = T) |> 
    rename(n_subjects = n)

# gets only those who remained in the subject, then counts by student
enrolments |> 
    filter(session == 202230, 
           is.na(withdraw_date) # exluding those that have withdrawn
           ) |> 
    count(id, sort = T) |> 
    rename(n_subjects = n)
```

We might also be interested in what course these students are in. The toy data is primarily for the *Bachelor of Nursing*, however joining directly to the *student_course* table can produce errors if a student has been in multiple courses.

``` {r}
enrolments |> 
    filter(session == 202230) |> 
    distinct(id, session) |> 
    inner_join(
        student_course |> 
            select(id, course, course_code, 
                   admit_session, last_registered_session))
```

We need to select the most appropriate course, based on the `admit_session` (when the student entered the course - however be warned this is sometimes later than their first subjects in the course!), and their `last_registered_session`. Then it might be necessary to ignore the `course_code` field as there can be multiple course codes for a single course.

``` {r}
students_in_202230 <- 
    enrolments |> 
    filter(session == 202230) |> 
    distinct(id, session) |> 
    inner_join(
        student_course |> 
            select(id, course, course_code, 
                   admit_session, last_registered_session)) |> 
    group_by(id, session) |>
    filter(
        # exclude any courses if already completed
        last_registered_session >= session | is.na(last_registered_session),
        # 
        admit_session <= session
        ) |> 
    distinct(id, session, course) |> 
    ungroup()

students_in_202230
```

