---
title: "Working-with-enrolments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working-with-enrolments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(retention.helpers)
```

## Overview

Enrolments come in two levels at Charles Sturt University; course and subject. Course enrolments refer to the students' enrolment in a particular program of study, such as a Bachelor of Nursing. These are stored in the `student_course` table. Subject level enrolments are stored in the `enrolments` table, and correspond to the enrolment records for units of study in a particular session. These should match to the `academic` table that stores the grades for each subject taken (if the student remained past census date). 

### Quirks

The `enrolments` tables also includes other pseudo-subjects, that are place holders for a Leave of Absence. These 'subjects' have the code **XLV0000** and should not be treated like a normal subject. Subjects beginning with the code **SSS...** might also be needed to be treated differently, as they are free subjects and exhibit very different behaviour from students. 

## Counting enrolments

For this example, we set up a toy version of the `student_course` and `enrolments` tables:

